Storm surge modelling in the Strait of Georgia

### 2014-12-13 AO_guides/manuscriptAO.tex
Started converting to AO style. Need to figure out citations!

\documentclass[pdftex,10pt]{article}
\usepackage{fullpage}
\usepackage{graphicx}
\usepackage{natbib}
\usepackage{amsmath}
\usepackage{setspace} \doublespacing

\title{Storm surge modelling in the Strait of Georgia}
\author{Nancy Soontiens, Susan Allen, Doug Latornell, Kate Le Sou\"{e}f, Idalia Machuca,
\\ Youyu Lu, Keith Thompson, Jean-Philippe Paquin, Vasily Korabel}
\date{\today}

\begin{document}
\maketitle

@@

@@

\begin{tabular}{|l |c c c c c | c c c c c|} 

@@

\end{tabular}
\caption{Evaluation of model performance for all hindcasts presented. Mean absolute error ($\bar{e}$), observed mean (subscript o), modelled mean (subscript m), $\gamma^2$, and the Willmott skill score ($WS$) are provided for water level predictions ($\eta$) and residuals ($\delta$). Statistics are calculated over the time period displayed in each simulation's corresponding figure (three or four days). For the Nov 2006 case, the statistics were calculated over a three day period. The average values were calculated over all of the hindcasts, excluding the Dec 2012 HRDPS simulation. }

@@

%\begin{table}[h]
%\centering 
%\begin{tabular}{|l |c c c c c | c c c c c|} 
%\hline 
%& \multicolumn{5}{|c|}{Water Level ($\eta$)}        & \multicolumn{5}{|c|}{Residual ($\delta$)} \\ 
%\hline 
%Location       & $\bar{e}$ (cm) & $\bar{\eta_{o}}$ (m) & $\bar{\eta_{m}}$ (m) & $\gamma^2$ & $WS$   & $\bar{e}$ (cm) & $\bar{\delta_{o}}$ (m) & $\bar{\delta_{m}}$ (m) & $\gamma^2$ & $WS$ \\
%\hline 
%Point Atkinson &  9.8           &  3.42                & 3.43                 &   0.034    & 0.991  &  7.7           &  0.290                 & 0.247                  &  0.128     & 0.950 \\
%Victoria       &  9.1           &  2.16                & 2.11                 &   0.070    & 0.978  &  6.2           &  0.222                 & 0.202                  &  0.171     & 0.946 \\
%Patricia Bay   & 10.0           &  2.57                & 2.53                 &   0.049    & 0.985  &  6.6           &  0.235                 & 0.203                  &  0.144     & 0.947 \\
%Campbell River & 10.7           &  3.20                & 3.26                 &   0.037    & 0.989  &  6.2           &  0.254                 & 0.236                  &  0.140     & 0.952 \\
%\hline 
%\end{tabular}
%\caption{Evaluation of model performance for the Nov 2006 hindcast. Statistics were calculated over the period Nov 14, 2006 to Nov 17, 2006.}
%\label{tab:nov2006stat} 
%\end{table}

@@

%\begin{table}[h]
%\centering 
%\begin{tabular}{|l |c c c c c | c c c c c|} 
%\hline 
%& \multicolumn{5}{|c|}{Water Level ($\eta$)}        & \multicolumn{5}{|c|}{Residual ($\delta$)} \\ 
%\hline 
%Location       & $\bar{e}$ (cm) & $\bar{\eta_{o}}$ (m) & $\bar{\eta_{m}}$ (m) & $\gamma^2$ & $WS$   & $\bar{e}$ (cm) & $\bar{\delta_{o}}$ (m) & $\bar{\delta_{m}}$ (m) & $\gamma^2$ & $WS$ \\
%\hline 
%Point Atkinson & 10.5           &  3.45                & 3.47                 &   0.025    & 0.994  & 11.1           &  0.212                 & 0.186                  &  0.275     & 0.891 \\
%Victoria       &  9.6           &  2.23                & 2.16                 &   0.042    & 0.984  &  7.9           &  0.190                 & 0.156                  &  0.223     & 0.911 \\
%Patricia Bay   &  9.2           &  2.63                & 2.58                 &   0.031    & 0.990  &  9.5           &  0.193                 & 0.149                  &  0.260     & 0.889 \\
%Campbell River & 10.6           &  3.23                & 3.28                 &   0.030    & 0.992  &  9.4           &  0.191                 & 0.163                  &  0.247     & 0.905 \\
%\hline 
%\end{tabular}
%\caption{Evaluation of model performance for the Dec 2006 hindcast. Statistics were calculated over the period Dec 14, 2006 to Dec 17, 2006.}
%\label{tab:dec2006stat} 
%\end{table}

@@

%\begin{table}[h]
%\centering 
%\begin{tabular}{|l |c c c c c | c c c c c|} 
%\hline 
%& \multicolumn{5}{|c|}{Water Level ($\eta$)}        & \multicolumn{5}{|c|}{Residual ($\delta$)} \\ 
%\hline 
%Location       & $\bar{e}$ (cm) & $\bar{\eta_{o}}$ (m) & $\bar{\eta_{m}}$ (m) & $\gamma^2$ & $WS$   & $\bar{e}$ (cm) & $\bar{\delta_{o}}$ (m) & $\bar{\delta_{m}}$ (m) & $\gamma^2$ & $WS$ \\
%\hline 
%Point Atkinson & 14.0           &  3.44                & 3.55                 &   0.012    & 0.995  &  6.1           &  0.313                 & 0.331                  &  0.237     & 0.914 \\
%Victoria       &  9.9           &  2.26                & 2.24                 &   0.023    & 0.994  &  4.8           &  0.323                 & 0.293                  &  0.245     & 0.899 \\
%Patricia Bay   & 11.2           &  2.60                & 2.65                 &   0.020    & 0.993  &  6.0           &  0.277                 & 0.290                  &  0.272     & 0.901 \\
%Campbell River & 13.5           &  3.22                & 3.27                 &   0.019    & 0.995  &  4.9           &  0.326                 & 0.314                  &  0.217     & 0.924 \\
%\hline 
%\end{tabular}
%\caption{Evaluation of model performance for the Nov 2009 hindcast. Statistics were generated over the period Nov 17, 2009 to Nov 22, 2009.}
%\label{tab:nov2009stat} 
%\end{table}

@@

%Where can I find a reference for the high res atmospheric model?

@@

\appendix
\section{Appendix}\label{sec:appendix}
This section details the procedure used to tune the tidal forcing for a good match with observations in the Strait of Georgia. Originally, the eight most important tidal constituents were extracted from a tide prediction tool called Webtide \citep{foreman2000webtide} at the western Juan de Fuca boundary. At the northern boundary,  M$_2$ and K$_1$ tidal harmonics for both elevation and currents were taken from \citet{thomson1980johnstone}, as well as elevation harmonics for O$_1$ and S$_2$. The remaining constituents were extrapolated from Webtide based on the ratio between S$_2$/M$_2$ amplitudes and phase differences for the semi-diurnal constituents and O$_1$/K$_1$ for the diurnal constituents. However, these forcing conditions did not result in a good match between observed and modelled tidal elevation harmonics in the Strait of Georgia, the main region of interest (although the matches in the Strait of Juan de Fuca were adequate). As such, tidal forcing was tuned, by first adjusting the western boundary M$_2$ and K$_1$ phase and K$_1$ amplitude for a good match in the Strait of Georgia (specifically at Point Atkinson, Gibsons Landing, Winchelsea and Halfmoon Bay). This tuning was performed on simulations that used only M$_2$ and K$_1$ forcing. Next, the M$_2$ and K$_1$ fluxes at the northern boundary were adjusted because the cross-sectional area of transect used by \citet{thomson1980johnstone} to measure tidal currents in Johnstone Strait is smaller than the cross-sectional area of our northern boundary. Finally, the remaining constituents were added, starting with O$_1$, then S$_2$, and then in pairs P$_1$/N$_2$ and Q$_1$/K$_2$. Final adjustments were made by slightly varying amplitude and phase, evaluating the changes to M$_2$, K$_1$ and the ratios S$_2$/M$_2$, N$_2$/M$_2$, K$_2$/M$_2$, O$_1$/K$_1$, P$_1$/K$_1$, Q$_1$/K$_1$ and then choosing new values to move closer to observations by linear interpolation.

@@

\begin{table}[h]
\centering 
\begin{tabular}{l l l l c c c c c c} 
\hline 
& & & & \multicolumn{3}{c}{M$_2$} & \multicolumn{3}{c}{K$_1$} \\ 
\hline 
            & Station name     & Longitude & Latitude & $R$ & $ \Delta \phi$ (deg) & $D$ (cm) & $R$ & $ \Delta \phi$ (deg) & $D$ (cm) \\
\hline 
1 & Port Renfrew & -124.43 & 48.55 & 1.16 & -9.4 & 16.63  & 1.03 & -4.94 & 4.17 \\
2 & Sheringham Point & -123.92 & 48.38 & 1.19 & -16.47 & 17.95  & 1 & -4.44 & 4.24 \\
3 & Pedder Bay & -123.55 & 48.33 & 0.96 & -27.39 & 15.91  & 0.99 & -2.41 & 2.66 \\
4 & Esquimalt & -123.43 & 48.43 & 0.91 & -23.09 & 14.40  & 0.99 & -1.04 & 1.21 \\
5 & Victoria & -123.37 & 48.42 & 0.89 & -17.77 & 11.62  & 1.04 & -1.83 & 3.22 \\
6 & Clover Point & -123.35 & 48.41 & 0.83 & -17.95 & 13.39  & 1.03 & -2.23 & 3.14 \\
7 & Finnerty Cove & -123.3 & 48.47 & 0.83 & -8.24 & 9.47  & 1.02 & -2.19 & 3.25 \\
8 & Fulford Harbour & -123.45 & 48.77 & 0.82 & -3.08 & 10.78  & 1 & -0.63 & 0.83 \\
9 & Tumbo Channel & -123.11 & 48.79 & 1.05 & 2.03 & 4.43  & 1.04 & 0.66 & 3.34 \\
10 & Patos Island & -122.97 & 48.78 & 1.03 & 0.26 & 2.32  & 1.04 & -1.5 & 3.92 \\
11 & Whaler Bay & -123.33 & 48.89 & 1 & -0.76 & 1.10  & 1.01 & -0.48 & 1.19 \\
12 & Tsawwassen & -123.13 & 48.99 & 1 & 0.92 & 1.31  & 1.01 & 0.19 & 1.09 \\
13 & Sandheads & -123.3 & 49.1 & 1.02 & -1.05 & 2.58  & 1.04 & -0.59 & 3.15 \\
14 & Point Grey & -123.27 & 49.25 & 0.98 & -3.31 & 5.83  & 0.96 & -0.68 & 3.63 \\
15 & Point Atkinson & -123.25 & 49.33 & 1.01 & -0.48 & 1.07  & 1 & -0.56 & 0.94 \\
16 & Gibsons Landing & -123.5 & 49.4 & 0.98 & 0.75 & 2.34  & 1 & 0.96 & 1.47 \\
17 & Winchelsea & -124.08 & 49.3 & 1.01 & -0.19 & 0.78  & 1.01 & 0.51 & 1.30 \\
18 & Halfmoon Bay & -123.92 & 49.52 & 1.01 & -0.15 & 0.83  & 1.01 & 0.15 & 0.72\\ 
19 & Irvines Landing & -124.05 & 49.63 & 1 & -0.32 & 0.59  & 1.01 & 0.24 & 1.12 \\
20 & Powell River & -124.55 & 49.86 & 1 & -1.72 & 3.05  & 1 & -0.01 & 0.43 \\
21 & Little River & -124.92 & 49.74 & 1.03 & 0.07 & 2.54  & 1.01 & 0.33 & 0.73\\ 
22 & Lund & -124.76 & 49.98 & 1.02 & -2.52 & 4.98  & 1.02 & -0.95 & 2.45 \\
23 & Twin Islets & -124.93 & 50.03 & 1.02 & -2.05 & 4.22  & 1.01 & -0.24 & 0.72\\ 
24 & Campbell River & -125.24 & 50.04 & 0.99 & 0.51 & 1.12  & 1 & 0.94 & 1.39 \\
25 & Maude Island E & -125.33 & 50.13 & 1.13 & -10.02 & 12.62  & 0.97 & -1.44 & 2.89\\ 
26 & Nymphe Cove & -125.36 & 50.13 & 1.02 & -5.43 & 6.04  & 0.97 & -0.29 & 2.02 \\
27 & Seymour Narrows & -125.34 & 50.13 & 0.63 & 31.31 & 53.40  & 1.14 & 10.91 & 17.28 \\
28 & Brown Bay & -125.37 & 50.16 & 0.98 & -5.63 & 9.26  & 0.98 & -0.04 & 1.17 \\
29 & Chatham Point & -125.44 & 50.33 & 1.09 & -5.49 & 12.18  & 0.98 & -3.18 & 3.85 \\
30 & Kelsey Bay & -125.96 & 50.39 & 1 & -1.69 & 3.50  & 0.98 & -0.39 & 1.00 \\
31 & Yorke Island & -125.97 & 50.44 & 0.99 & 1.59 & 3.32  & 1.01 & 1.23 & 1.37\\ 
\hline 
&   Mean Error (cm/deg) & & & 4.23  & 6.5 & 8.05   & 1.56 & 1.49 & 2.58 \\
&   RMS Error (cm/deg)  & & & 7.67  & 10.73 & 12.73 & 2.41 & 2.57 & 3.91 \\
\hline  
\end{tabular}
\caption{Comparisons of modelled and observed amplitudes and phases. $R$ is the ratio of modelled to observed amplitude. $\Delta \phi$ is the difference between modelled and observed phase. $D$ is the complex difference. The mean error and root mean square (RMS) error are calculated as the difference between the modelled quantity and observed quantity. }
\label{tab:comparison} 
\end{table}

\section{Acknowledgements}\label{sec:ack}
We would like to thank Michael Foreman, Diane Masson, Luc Fillion, Rich Pawlowicz, Mark Halverson, Kao-Shen Chung, and Pramod Thupaki for providing data used in model set up and evaluation. For providing guidance with stakeholder engagement, we also thank Stephanie Chang, Jackie Yip, and Shona van Zijil de jong. This project is funded by the Marine Environmental Observation Prediction and Response Network (MEOPAR), a Network of Centres of Excellence of Canada. This research was enabled in part by support provided by WestGrid (www.westgrid.ca) and Compute Canada Calcul Canada (www.computecanada.ca).

>>>>

% tATOguide.tex
% v1.2 released April 2014

\documentclass{tATO2e}

\usepackage{subfigure}% Support for small, `sub' figures and tables

\usepackage[longnamesfirst,sort]{natbib}% Citation support using natbib.sty
\bibpunct[, ]{(}{)}{;}{a}{,}{,}% Citation support using natbib.sty

%\usepackage[natbibapa]{apacite}% Citation support using apacite.sty. Commands using natbib.sty MUST be deactivated first!

\begin{document}

%\jvol{00} \jnum{00} \jyear{2014} \jmonth{April}

%\articletype{GUIDE}% NOT REQUIRED IN A RESEARCH ARTICLE

\title{Storm surge modelling in the Strait of Georgia}

\author{Nancy Soontiens$^{1}$$^{\ast}$\thanks{$^\ast$Corresponding author's email: nsoontie@eos.ubc.ca
\vspace{6pt}}, Susan Allen$^{1}$,  Doug Latornell$^{1}$,  Kate Le Sou\"{e}f$^{1}$, Idalia Machuca$^{1}$ \\ 
Youyu Lu$^{2}$, Keith Thompson$^{3}$, Jean-Philippe Paquin$^{3}$, Vasily Korabel$^{3}$
\\\vspace{6pt}  $^{1}${\em{University of British Columbia}};
$^{2}${\em{Bedford Insitute of Oceanography}}; 
$^{3}${\em{Dalhousie University}}\\\received{v1.2 released April 2014} }

\maketitle

@@

\begin{keywords}storm surge; Strait of Georgia; Salish Sea; numerical modelling; spatial variation
\end{keywords}

@@

\tbl{Evaluation of model performance for all hindcasts presented. Mean absolute error ($\bar{e}$), observed mean (subscript o), modelled mean (subscript m), $\gamma^2$, and the Willmott skill score ($WS$) are provided for water level predictions ($\eta$) and residuals ($\delta$). Statistics are calculated over the time period displayed in each simulation's corresponding figure (three or four days). For the Nov 2006 case, the statistics were calculated over a three day period. The average values were calculated over all of the hindcasts, excluding the Dec 2012 HRDPS simulation. } 
{\begin{tabular}{|l |c c c c c | c c c c c|} 

@@

\end{tabular}}

@@

@@

@@

@@

@@

\appendices
\section{Appendix}\label{sec:appendix}
This section details the procedure used to tune the tidal forcing for a good match with observations in the Strait of Georgia. Originally, the eight most important tidal constituents were extracted from a tide prediction tool called Webtide \citep{foreman2000webtide} at the western Juan de Fuca boundary. At the northern boundary,  M$_2$ and K$_1$ tidal harmonics for both elevation and currents were taken from \citet{thomson1980johnstone}, as well as elevation harmonics for O$_1$ and S$_2$. The remaining constituents were extrapolated from Webtide based on the ratio between S$_2$/M$_2$ amplitudes and phase differences for the semi-diurnal constituents and O$_1$/K$_1$ for the diurnal constituents. However, these forcing conditions did not result in a good match between observed and modelled tidal elevation harmonics in the Strait of Georgia, the main region of interest (although the matches in the Strait of Juan de Fuca were adequate). As such, tidal forcing was tuned, by first adjusting the western boundary M$_2$ and K$_1$ phase and K$_1$ amplitude for a good match in the Strait of Georgia (specifically at Point Atkinson, Gibsons Landing, Winchelsea and Halfmoon Bay). This tuning was performed on simulations that used only M$_2$ and K$_1$ forcing. Next, the M$_2$ and K$_1$ fluxes at the northern boundary were adjusted because the cross-sectional area of transect used by \citet{thomson1980johnstone} to measure tidal currents in Johnstone Strait is smaller than the cross-sectional area of our northern boundary. Finally, the remaining constituents were added, starting with O$_1$, then S$_2$, and then in pairs P$_1$/N$_2$ and Q$_1$/K$_2$. Final adjustments were made by slightly varying amplitude and phase, evaluating the changes to M$_2$, K$_1$ and the ratios S$_2$/M$_2$, N$_2$/M$_2$, K$_2$/M$_2$, O$_1$/K$_1$, P$_1$/K$_1$, Q$_1$/K$_1$ and then choosing new values to move closer to observations by linear interpolation.

@@

\begin{table}[h]
\centering 
\tbl{Comparisons of modelled and observed amplitudes and phases. $R$ is the ratio of modelled to observed amplitude. $\Delta \phi$ is the difference between modelled and observed phase. $D$ is the complex difference. The mean error and root mean square (RMS) error are calculated as the difference between the modelled quantity and observed quantity. }
{\begin{tabular}{l l l l c c c c c c} 
\hline 
& & & & \multicolumn{3}{c}{M$_2$} & \multicolumn{3}{c}{K$_1$} \\ 
\hline 
            & Station name     & Longitude & Latitude & $R$ & $ \Delta \phi$ (deg) & $D$ (cm) & $R$ & $ \Delta \phi$ (deg) & $D$ (cm) \\
\hline 
1 & Port Renfrew & -124.43 & 48.55 & 1.16 & -9.4 & 16.63  & 1.03 & -4.94 & 4.17 \\
2 & Sheringham Point & -123.92 & 48.38 & 1.19 & -16.47 & 17.95  & 1 & -4.44 & 4.24 \\
3 & Pedder Bay & -123.55 & 48.33 & 0.96 & -27.39 & 15.91  & 0.99 & -2.41 & 2.66 \\
4 & Esquimalt & -123.43 & 48.43 & 0.91 & -23.09 & 14.40  & 0.99 & -1.04 & 1.21 \\
5 & Victoria & -123.37 & 48.42 & 0.89 & -17.77 & 11.62  & 1.04 & -1.83 & 3.22 \\
6 & Clover Point & -123.35 & 48.41 & 0.83 & -17.95 & 13.39  & 1.03 & -2.23 & 3.14 \\
7 & Finnerty Cove & -123.3 & 48.47 & 0.83 & -8.24 & 9.47  & 1.02 & -2.19 & 3.25 \\
8 & Fulford Harbour & -123.45 & 48.77 & 0.82 & -3.08 & 10.78  & 1 & -0.63 & 0.83 \\
9 & Tumbo Channel & -123.11 & 48.79 & 1.05 & 2.03 & 4.43  & 1.04 & 0.66 & 3.34 \\
10 & Patos Island & -122.97 & 48.78 & 1.03 & 0.26 & 2.32  & 1.04 & -1.5 & 3.92 \\
11 & Whaler Bay & -123.33 & 48.89 & 1 & -0.76 & 1.10  & 1.01 & -0.48 & 1.19 \\
12 & Tsawwassen & -123.13 & 48.99 & 1 & 0.92 & 1.31  & 1.01 & 0.19 & 1.09 \\
13 & Sandheads & -123.3 & 49.1 & 1.02 & -1.05 & 2.58  & 1.04 & -0.59 & 3.15 \\
14 & Point Grey & -123.27 & 49.25 & 0.98 & -3.31 & 5.83  & 0.96 & -0.68 & 3.63 \\
15 & Point Atkinson & -123.25 & 49.33 & 1.01 & -0.48 & 1.07  & 1 & -0.56 & 0.94 \\
16 & Gibsons Landing & -123.5 & 49.4 & 0.98 & 0.75 & 2.34  & 1 & 0.96 & 1.47 \\
17 & Winchelsea & -124.08 & 49.3 & 1.01 & -0.19 & 0.78  & 1.01 & 0.51 & 1.30 \\
18 & Halfmoon Bay & -123.92 & 49.52 & 1.01 & -0.15 & 0.83  & 1.01 & 0.15 & 0.72\\ 
19 & Irvines Landing & -124.05 & 49.63 & 1 & -0.32 & 0.59  & 1.01 & 0.24 & 1.12 \\
20 & Powell River & -124.55 & 49.86 & 1 & -1.72 & 3.05  & 1 & -0.01 & 0.43 \\
21 & Little River & -124.92 & 49.74 & 1.03 & 0.07 & 2.54  & 1.01 & 0.33 & 0.73\\ 
22 & Lund & -124.76 & 49.98 & 1.02 & -2.52 & 4.98  & 1.02 & -0.95 & 2.45 \\
23 & Twin Islets & -124.93 & 50.03 & 1.02 & -2.05 & 4.22  & 1.01 & -0.24 & 0.72\\ 
24 & Campbell River & -125.24 & 50.04 & 0.99 & 0.51 & 1.12  & 1 & 0.94 & 1.39 \\
25 & Maude Island E & -125.33 & 50.13 & 1.13 & -10.02 & 12.62  & 0.97 & -1.44 & 2.89\\ 
26 & Nymphe Cove & -125.36 & 50.13 & 1.02 & -5.43 & 6.04  & 0.97 & -0.29 & 2.02 \\
27 & Seymour Narrows & -125.34 & 50.13 & 0.63 & 31.31 & 53.40  & 1.14 & 10.91 & 17.28 \\
28 & Brown Bay & -125.37 & 50.16 & 0.98 & -5.63 & 9.26  & 0.98 & -0.04 & 1.17 \\
29 & Chatham Point & -125.44 & 50.33 & 1.09 & -5.49 & 12.18  & 0.98 & -3.18 & 3.85 \\
30 & Kelsey Bay & -125.96 & 50.39 & 1 & -1.69 & 3.50  & 0.98 & -0.39 & 1.00 \\
31 & Yorke Island & -125.97 & 50.44 & 0.99 & 1.59 & 3.32  & 1.01 & 1.23 & 1.37\\ 
\hline 
&   Mean Error (cm/deg) & & & 4.23  & 6.5 & 8.05   & 1.56 & 1.49 & 2.58 \\
&   RMS Error (cm/deg)  & & & 7.67  & 10.73 & 12.73 & 2.41 & 2.57 & 3.91 \\
\hline  
\end{tabular}}
\label{tab:comparison} 
\end{table}

\section*{Acknowledgements}\label{sec:ack}
We would like to thank Michael Foreman, Diane Masson, Luc Fillion, Rich Pawlowicz, Mark Halverson, Kao-Shen Chung, and Pramod Thupaki for providing data used in model set up and evaluation. For providing guidance with stakeholder engagement, we also thank Stephanie Chang, Jackie Yip, and Shona van Zijil de jong. This project is funded by the Marine Environmental Observation Prediction and Response Network (MEOPAR), a Network of Centres of Excellence of Canada. This research was enabled in part by support provided by WestGrid (www.westgrid.ca) and Compute Canada Calcul Canada (www.computecanada.ca).
